---
title: "Diversity and ecology of oomycetes associated wiht soybean seedlings"
author: "J. Alejandro Rojas"
date: "April 19, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r custom_functions, echo=FALSE, message=FALSE}
# ipak function: install and load multiple R packages.
# check to see if packages are installed. Install them if they are not, 
#then load them into the R session.
# Source: https://gist.github.com/stevenworthington/3178163
ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
sapply(pkg, require, character.only = TRUE)
}

#Loading packages
packages <- c("phyloseq","ggplot2","RColorBrewer","plyr","dplyr",
              "tidyr","knitr","magrittr","ape","vegan", "ggtree", 
              "cowplot", "ggrepel", "gtable", "gridExtra")
ipak(packages)

#Standard error function
std <- function(x) sd(x)/sqrt(length(x))
CVcoeff <- function(x) (sd(x)/mean(x))*100

#Trimming leading or trailing space
trim <- function(x) gsub("^\\s+|\\s+$", "", x)

#RDP phyloseq extraction to vegan
veganotu <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  OTU <- as(OTU, "matrix")
  return(OTU)
}

vegansam <- function(physeq) {
  sam <- sample_data(physeq)
  # i <- sapply(sam, is.numeric)
  # j <- sapply(sam, is.character)
  # k <- sapply(sam, is.factor)
  # sam <- as.matrix(sam)
  sam <- data.frame(sam)
  # sam[i] <- lapply(sam[i], as.numeric)
  # sam[j] <- lapply(sam[j], as.character)
  # sam[k] <- lapply(sam[k], as.factor)
  return(sam)
}
```

##Read files for OTU analyses
```{r importing_files}
#Import BIOM, sample data and tree files
OTU_file <- "../data/clean/Seqs_11_12.OTU.biom"
Metadata_file <- "../data/clean/env_metadata.csv"

Oom_OTU <- import_biom(OTU_file)

#Oom_tree <- read_tree(Tree_file)
Oom_data <- import_qiime_sample_data(Metadata_file)

#Covert year to factor
Oom_data$Year <- as.factor(Oom_data$Year)

#Rename tax ranks to actual names
colnames(tax_table(Oom_OTU)) <- c("Phylum","Class","Order",
                                  "Family","Genus","Clade","Species")

#Merge phyloseq objects
Oom_biom <- merge_phyloseq(Oom_OTU, Oom_data) 
```

##Read files for phylotype analyses
```{r phylotype_files}
#Import files
otu_phylo <- read.csv("../data/clean/OTU.phylotype.txt", sep = "\t", 
                      row.names = 1, header = TRUE)
taxa_phylo <- read.csv("../data/clean/Taxa.phylotype.txt", sep = "\t", 
                       row.names = 1, header = TRUE)

#Transform to phyloseq objects
otu_phylo <- otu_table(otu_phylo, taxa_are_rows = TRUE)
taxa_phylo <- tax_table(as.matrix(taxa_phylo))

#Phyloseq object
Oom_phylo <- phyloseq(otu_phylo, taxa_phylo)
Oom_phylo <- merge_phyloseq(Oom_phylo, Oom_data)
```

##Richness analyses
```{r Richness_analysis, fig.align='center', fig.width=11, fig.height=7}
#Plotting richness
plot_richness(Oom_biom, "St_Yr", 
              measures = c("InvSimpson","Shannon","Chao1"), color = "Year")

#Table richness
Tb_richness <- estimate_richness(Oom_biom, 
                                 split = TRUE, 
                                 c("Observed", "Shannon", "Simpson")) %>% 
  add_rownames(var = "sample") %>%
  mutate(Evenness = Shannon/log(Observed))

samp_size <- colSums(otu_table(Oom_biom))
samp_size <- data.frame(samp_size) %>% add_rownames(var = "sample")

smp_state <- data.frame(sample_data(Oom_biom)[,1:5]) %>% 
  add_rownames(var = "sample")

Tb_richness_final <- left_join(Tb_richness, samp_size, by = "sample") %>%
  left_join(smp_state, by = "sample") %>%
  filter(Observed > 1) %>%
  ddply(c("St_Yr"), summarise, 
        N = length(sample),
        Isolates = sum(samp_size),
        mean.Observed = mean(Observed),
        sd.Observed = sd(Observed, na.rm = TRUE),
        mean.Shannon = mean(Shannon),
        sd.Shannon = sd(Shannon, na.rm = TRUE),
        mean.Simpson = mean(Simpson),
        sd.Simpson = sd(Simpson, na.rm = TRUE),
        mean.Evenness = mean(Shannon/log(Observed)),
        sd.Evenness = sd(Shannon/log(Observed), na.rm = TRUE))

kable(Tb_richness_final, digits = 3)
```

```{r plots_phylotype, fig.align='center', fig.width=10, fig.height=7}
#Summarizing by state
Oom_phylo_state <- tax_glom(Oom_phylo, taxrank = "Clade")
Oom_phylo_state <- merge_samples(Oom_phylo_state, "St_Yr")

#Transform counts for plot
#Oom_phylo_state <- transform_sample_counts(Oom_phylo_state, 
#                                           function(x) 100 * x/sum(x))
State_Year <- psmelt(Oom_phylo_state)

#Color scale
pal <- colorRampPalette(brewer.pal(12, "Paired"))

#Reorder factor
Clade_factor <- State_Year %>% group_by(Clade) %>% dplyr::summarise(sum(Abundance))
Clade_factor <- Clade_factor[order(-Clade_factor$`sum(Abundance)`),]
Clade_factor <- Clade_factor$Clade
State_Year$Clade <- factor(State_Year$Clade, levels = Clade_factor)
levels(State_Year$Clade)

data_state <- dplyr::select(State_Year, Sample, Abundance, Clade)
data_state <- data_state[with(data_state, order(Clade, as.numeric(Clade))),]

#Plot
(ByState <- ggplot(data = data_state, aes(Sample, Abundance, fill = Clade)) +
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
  scale_fill_manual(values = pal(18)) + 
   theme(text = element_text(size = 15)) + theme_gray())
```


```{r plots_otu, echo=FALSE, fig.align='center', fig.width=10, fig.height=7}
#Summarizing by State
Oom_biom_state <- tax_glom(Oom_biom, taxrank = "Clade")
Oom_biom_state <- merge_samples(Oom_biom_state, "St_Yr")

classified <- as.vector(substr(tax_table(Oom_biom_state)[,"Clade"], 
                               0, 7) != "unclass")
Oom_biom_state <- prune_taxa(classified, Oom_biom_state)

#Color scale
pal <- colorRampPalette(brewer.pal(12, "Paired"))

#Data_OTU abundance plot
State_OTU <- psmelt(Oom_biom_state)
clf_otu <- as.vector((State_Year %>% 
                        group_by(Clade) %>% 
                        summarise(sum(Abundance)))$Clade)
State_OTU$Clade <- factor(State_OTU$Clade, levels = clf_otu)
State_OTU <-  dplyr::select(State_OTU, Sample, Abundance, Clade)
State_OTU <- State_OTU[with(State_OTU, order(Clade, as.numeric(Clade))),]

(State_abd_otu <- ggplot(State_OTU, aes(Sample, Abundance, fill = Clade)) + 
  geom_bar(stat = "identity", position = position_fill()) + coord_flip() +
  scale_fill_manual(values = pal(16)) + theme_gray())
```

## Rarefaction curves
```{r rarefaction_curvers, warning=FALSE, fig.align='center', fig.width=12, fig.height=7}
## DATA ##

psdata <- merge_samples(Oom_phylo, "St_Yr")
psdata
sample_sums(psdata)

### Calculate alpha diversity ###
set.seed(42)

calculate_rarefaction_curves <- function(psdata, measures, depths) {
  require('plyr') # ldply
  require('reshape2') # melt

  estimate_rarified_richness <- function(psdata, measures, depth) {
    if(max(sample_sums(psdata)) < depth) return()
    psdata <- prune_samples(sample_sums(psdata) >= depth, psdata)

    rarified_psdata <- rarefy_even_depth(psdata, depth, verbose = FALSE)

    alpha_diversity <- estimate_richness(rarified_psdata, measures = measures)

    # as.matrix forces the use of melt.array, which includes the Sample names (rownames)
    molten_alpha_diversity <- melt(as.matrix(alpha_diversity), varnames = c('Sample', 'Measure'), value.name = 'Alpha_diversity')

    molten_alpha_diversity
  }

  names(depths) <- depths # this enables automatic addition of the Depth to the output by ldply
  rarefaction_curve_data <- ldply(depths, estimate_rarified_richness, psdata = psdata, measures = measures, .id = 'Depth', .progress = ifelse(interactive(), 'text', 'none'))

  # convert Depth from factor to numeric
  rarefaction_curve_data$Depth <- as.numeric(levels(rarefaction_curve_data$Depth))[rarefaction_curve_data$Depth]

  rarefaction_curve_data
}

rarefaction_curve_data <- calculate_rarefaction_curves(psdata, c('Observed', 'Shannon'), rep(c(1, 5, 10, 1:20 * 10), each = 10))
summary(rarefaction_curve_data)


### Summarize alpha diversity ###
rarefaction_curve_data_summary <- ddply(rarefaction_curve_data, c('Depth', 'Sample', 'Measure'), summarise, Alpha_diversity_mean = mean(Alpha_diversity), Alpha_diversity_sd = sd(Alpha_diversity))

### Add sample data ###
smp_dt <- as.data.frame(sample_data(psdata))
row.names(smp_dt) <- gsub(pattern = " ",replacement = ".", x = row.names(smp_dt))
rarefaction_curve_data_summary_verbose <- merge(rarefaction_curve_data_summary, smp_dt,
                                                by.x = 'Sample', by.y = 'row.names')

### plot ###
ggplot(
  data = rarefaction_curve_data_summary_verbose,
  mapping = aes(
    x = Depth,
    y = Alpha_diversity_mean,
    ymin = Alpha_diversity_mean - Alpha_diversity_sd,
    ymax = Alpha_diversity_mean + Alpha_diversity_sd,
    colour = State2,
    group = Sample)) + 
  geom_smooth() + 
  #geom_pointrange() + 
  facet_wrap(facets = ~ Measure, scales = 'free_y') + 
  theme_gray()
```


## Latitude/Longitude gradient
```{r latitude_vs_diversity, fig.align='center', fig.width=7, fig.height=4}
#Richness data
richness2 <- estimate_richness(Oom_biom, 
                                 split = TRUE, 
                                 c("Observed", "Shannon", "Simpson", "Chao1", "InvSimpson")) %>% 
  add_rownames(var = "sample")


samp_size <- colSums(otu_table(Oom_biom))
samp_size <- data.frame(samp_size) %>% add_rownames(var = "sample")

#Sample data
smp_state2 <- data.frame(sample_data(Oom_biom)[,c(1:5,8,9)]) %>% 
  add_rownames(var = "sample")
smp_state2 <- left_join(smp_state2, samp_size, by = "sample")

#Merging tables together
lt_rch_data <- left_join(smp_state2, richness2, by = "sample")
lt_rch_data <- lt_rch_data[!is.na(lt_rch_data[,7]),]
lt_rch_data <- lt_rch_data[(lt_rch_data[,9]) >= 10,]

#Correlation
cor.test(lt_rch_data$Lat, lt_rch_data$Observed, method = "spearman")
cor.test(lt_rch_data$Lat, lt_rch_data$Simpson, method = "spearman")
cor.test(lt_rch_data$Lat, lt_rch_data$Shannon, method = "spearman")
cor.test(lt_rch_data$Long, lt_rch_data$Shannon, method = "spearman")

#Plot
(Obs_plot <- ggplot(lt_rch_data, aes(x = Lat, y = Observed)) + 
  geom_point(size = 2, alpha = 0.5) + geom_smooth(method = lm) + theme_gray() +
  labs(x = "Latitude", y = "Observed OTUs") +
  annotate("text", x=45, y=0.4, label = "p-value = 0.0813\nrho = 0.189", 
           fontface = "bold", hjust = 0))
  

(Obs_plot2 <- ggplot(lt_rch_data, aes(x = Lat, y = Shannon)) + 
  geom_point(size = 2, alpha = 0.5) + geom_smooth(method = lm) + theme_gray() +
  labs(x = "Latitude", y = "Shannon diversity index") +
  annotate("text", x=45, y=0.4, label = "p-value = 0.016\nrho = 0.258", 
           fontface = "bold", hjust = 0))

(Obs_plot3 <- ggplot(lt_rch_data, aes(x = Long, y = Shannon)) + 
  geom_point(size = 2, alpha = 0.5) + geom_smooth(method = lm) + theme_gray() +
  labs(x = "Longitude", y = "Shannon diversity index") +
  annotate("text", x=-88, y=0.4, label = "p-value = 0.037\nrho = -0.224", 
           fontface = "bold", hjust = 0))

#grid.arrange(Obs_plot2, Obs_plot3)
```

##Cluster analysis - OTU

```{r cluster_analysis_otu, fig.align='center', fig.width=8, fig.height=4}
#Phyloseq otu table to vegan otu table
Oom_st <- prune_samples(!(grepl('MICO|ONSO', sample_names(Oom_biom))), Oom_biom)
Oom_st <- merge_samples(Oom_biom, "St_Yr")
tb_otu <- veganotu(Oom_st)
tb_otu <- tb_otu[row.names(tb_otu) != "Ontario 2012",]

#Phyloseq sample to vegan sample table
tb_sample <- vegansam(Oom_st)

#Relative abundance and bray-curtis distance
tb_otu <- decostand(tb_otu, method = "total")
tb_otu.bc <- vegdist(tb_otu, method = "bray")
tb_otu.pa.bc <- vegdist(tb_otu, method = "bray", binary = TRUE)

tb.otu.cl <- as.phylo(hclust(tb_otu.bc, method = "ward.D2"))
tb.otu.pa <- as.phylo(hclust(tb_otu.pa.bc, method = "ward.D2"))

cls <- list(c1 = c("N Dakota 2011", "N Dakota 2012", 
                   "Minnesota 2011", "Minnesota 2012"),
            c2 = c("S Dakota 2011", "S Dakota 2012", "Iowa 2011", 
                   "Iowa 2012", "Nebraska 2011"),
            c3 = c("Wisconsin 2011", "Illinois 2011", "Illinois 2012", 
                   "Indiana 2011", "Indiana 2012", "Michigan 2011", 
                   "Michigan 2012"),
            c4 = c("Kansas 2011", "Kansas 2012", "Arkansas 2011", 
                   "Arkansas 2012"))

tb.otu.cl <- groupOTU(tb.otu.cl, cls)
tb.otu.pa <- groupOTU(tb.otu.pa, cls)

t1.otu <- ggtree(tb.otu.cl, branch.length = 'branch.length') + 
  #geom_text(aes(label = label, hjust = -0.05)) + 
  ggplot2::xlim(0,1) + geom_tiplab(aes(color = group, label = label)) +
  scale_color_brewer(type = "div", palette = "Set1")

ggtree(tb.otu.pa) + geom_text(aes(label = label, hjust = -0.05)) + ggplot2::xlim(0,0.6)

t2.otu <- ggtree(tb.otu.pa, branch.length = 'branch.length') + 
  #geom_text(aes(label = label, hjust = -0.05)) + 
  ggplot2::xlim(0,0.6) + geom_tiplab(aes(color = group, label = label)) +
  scale_color_brewer(type = "div", palette = "Set1")

gridExtra::grid.arrange(t1.otu, t2.otu, ncol = 2)
```


##Cluster analysis - phylotype

```{r cluster_analysis_phylo, fig.align='center', fig.width=8, fig.height=4}
#Phyloseq otu table to vegan otu table
Oom_ph.1 <- prune_samples(!(grepl('MICO|ONSO', sample_names(Oom_phylo))), Oom_phylo)
Oom_ph <- merge_samples(Oom_ph.1, "St_Yr")
tb_ph <- veganotu(Oom_ph)
tb_ph <- tb_ph[row.names(tb_ph) != "Ontario 2012",]

#Phyloseq sample to vegan sample table
tb_sp_ph <- vegansam(Oom_ph)

#Relative abundance and bray-curtis distance
tb_ph <- decostand(tb_ph, method = "total")
tb_ph.bc <- vegdist(tb_ph, method = "bray")
tb_ph.pa.bc <- vegdist(tb_ph, method = "jaccard", binary = TRUE)

tb.ph.cl <- as.phylo(hclust(tb_ph.bc))
tb.ph.pa <- as.phylo(hclust(tb_ph.pa.bc))

tb.ph.cl <- groupOTU(tb.ph.cl, cls)
tb.ph.pa <- groupOTU(tb.ph.pa, cls)


ggtree(tb.ph.cl, layout="rectangular") + 
  geom_text(aes(label = label, hjust = -0.05)) + ggplot2::xlim(0,0.6) + theme_tree2()

t1.ph <- ggtree(tb.ph.cl, branch.length = 'branch.length') + 
  #geom_text(aes(label = label, hjust = -0.05)) + 
  ggplot2::xlim(0,0.6) + geom_tiplab(aes(label = label)) +
  scale_color_brewer(type = "div", palette = "Set1") + xlab("Height") + ylab("Cluster dendogram")

ggtree(tb.ph.pa) + 
  geom_text(aes(label = label, hjust = -0.05)) + ggplot2::xlim(0,0.6)

t2.ph <- ggtree(tb.ph.pa, branch.length = 'branch.length') + 
  #geom_text(aes(label = label, hjust = -0.05)) + 
  ggplot2::xlim(0,0.6) + geom_tiplab(aes(color = group, label = label)) +
  scale_color_brewer(type = "div", palette = "Set1")

#gridExtra::grid.arrange(t1.ph, t2.ph, ncol = 2)
#gridExtra::grid.arrange(t1.otu, t2.otu, t1.ph, t2.ph, ncol = 2)
```

```{r, fig.align='center', fig.width=8, fig.height=10}
plot_grid(t1.ph, Obs_plot2, labels = c("A", "B"), ncol = 1,
          align = "h"
          #rel_widths = c(0.85,1,1), rel_heights = c(0.5,1,1) 
          )

(plot_grid(t1.ph, Obs_plot2, Obs_plot3, labels = c("A", "B", "C"), ncol = 1,
          align = "h"
          #rel_widths = c(0.85,1,1), rel_heights = c(0.5,1,1) 
          ))

```


## Analysis of similarity (ANOSIM) for different parameters

```{r anosim_test}
Oom_grp <- get_variable(Oom_biom, "State2")
Oom_st_ano <- anosim(distance(Oom_biom, "bray"), Oom_grp)

Oom_biom_lt <- prune_samples(!is.na(Oom_biom@sam_data$Lat), Oom_biom)

Lat_grp <- cut(get_variable(Oom_biom_lt, "Lat"), c(32,42,50))
(Oom_lt_ano <- anosim(distance(Oom_biom_lt, "bray"), Lat_grp))

Long_grp <- cut(get_variable(Oom_biom_lt, "Long"), c(-80,-95,-110))
(Oom_lg_ano <- anosim(distance(Oom_biom_lt, "bray"), Long_grp))

Yr_grp <- get_variable(Oom_biom, "Year")
Oom_yr_ano <- anosim(distance(Oom_biom, "bray"), Yr_grp)

StYr_grp <- get_variable(Oom_biom, "St_Yr")
Oom_styr_ano <- anosim(distance(Oom_biom, "bray"), StYr_grp)
```

## ADONIS for different parameters
```{r adonis_test}
df <- as(sample_data(Oom_biom), "data.frame")
d <- distance(Oom_biom, "bray")

Oom_adonis <- adonis(d ~ State2 + Year + State2*Year, df)

kable(Oom_adonis$aov.tab, digits = 3, 
      caption = "__Table 2.__ Comparison of community structure (beta diversity)\
      using Bray-curtis distance by State and year.")
```

##Ordination analysis
```{r Ordination_plot}
Oom_biom2 <- prune_samples(!is.na(Oom_biom@sam_data$Lat), Oom_biom)

colors2 <- c("#77C7C6",
"#C37F3B",
"#869BCF",
"#7DD54E",
"#C67BB7",
"#CDC84A",
"#CC6569",
"#83D693",
"#7A7678",
"#698547",
"#D3C1A7")

Oom_biom_ord <- ordinate(Oom_biom2, "PCoA", "bray")
ord_plot <- plot_ordination(Oom_biom2, Oom_biom_ord, color = "State2", shape = "Year")
(ord_plot.f <- ord_plot + geom_point(size = 4, alpha = 0.7) + 
  scale_colour_manual(values = colors2) +
  theme_light() + labs(color = "State", shape = "Year"))
```

## Environmental/Edaphic factor analysis

```{r envfit_analysis}
## Environment fit analysis
bray.pcoa <- ordinate(Oom_biom2, method = "PCoA", "bray")
env <- as.data.frame(Oom_biom2@sam_data)

Oom_env <- envfit(bray.pcoa$vectors, env, permutations = 999)

fit_data <- as.data.frame(scores(Oom_env, display = "vectors")) %>%
  add_rownames(var = "Env.var") %>%
  bind_cols(data.frame(Oom_env$vectors$r, Oom_env$vectors$pvals)) %>%
  rename(R2 = Oom_env.vectors.r, P.value = Oom_env.vectors.pvals) %>%
  arrange(P.value)
  
## Supplementary material version

kable(fit_data, digits = 3, caption = "__Supp. table 1.__ Significance and correlation\
of vectors fitted into PCoA ordination of oomycete communities associated with\
soybean seedlings")

## Reduced version

#kable(fit_data[fit_data$P.value < 0.05,], digits = 3, caption = "__Table 3.__ Significance and correlation\
#of vectors fitted into PCoA ordination of oomycete communities associated with\
#soybean seedlings")

```

### Results ordination and environmental data

```{r Ord_envfit_plot, fig.align='center', fig.width=10, fig.height=6}
## Vectors for plot
fit_reduced <- fit_data[fit_data$P.value < 0.05,] 

fit_plot <- as.data.frame(scores(Oom_env, display = "vectors")) %>%
  add_rownames(var = "Env.var") %>%
  inner_join(fit_reduced, by = "Env.var") %>%
  arrange(P.value) %>%
  slice(c(10,1,5,19,18,8,20,22,2,17,15,12,21,6,23))

fit_plot$Env.var2 <- c("Latitude", "Longitude", "Precip. Season","CEC", "Clay (%)", 
                       "Bulk density", "Soil pH", "Water content", 
                       "Annual Total Precip.", "Max. Temp 30yr", 
                       "Mean Temp 30yr", "Min. Temp 30yr",
                       "Annual Min. Temp","Precip. 30yr", "Min. Temp Season")

## paper version
kable(fit_plot, digits = 3, caption = "__Table 3.__ Significant factors using\
      ‘envfit’ function from vegan that affect oomycete community associated\
      with soybean seedlings.")

ord_plot.data <- plot_ordination(Oom_biom2, Oom_biom_ord, 
                            color = "State2", shape = "Year", justDF = TRUE)

(ord.plot.env <- ggplot(data = ord_plot.data, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(color=State2, shape=Year), size = 4, alpha = 0.7) + 
  #scale_color_brewer(type = "div", palette ="Spectral") +
  labs(color = "State", shape = "Year", x = "PCoA 1 [12.1%]", y = "PCoA 2 [9.4%]") +
  scale_colour_manual(values = colors2) +
  geom_segment(data = fit_plot, aes(x = 0, xend = Axis.1.x, y = 0, yend = Axis.2.x), 
               arrow = arrow(length = unit(0.1,"cm")), color = "black", size = 0.8) + 
  geom_label_repel(data = fit_plot, aes(x = Axis.1.x, y = Axis.2.x, label = Env.var2), 
            size = 3, force = 1) + #facet_wrap(~Year) +
  theme_gray())
```

## Correlation of environmental parameters with PCoA axes

```{r factor_correlation, fig.width=14, fig.height= 10}
#Season precipitation correlation with axis
P.cor <- cor.test(ord_plot.data[,"Axis.1"], log10(ord_plot.data[,"Precip_AMJ"]), method = "spearman")

Tmin.cor <- cor.test(ord_plot.data[,"Axis.1"], ord_plot.data[,"Tmin_AMJ"], method = "spearman")

Clay.cor <- cor.test(ord_plot.data[,"Axis.2"], ord_plot.data[,"Clay"], method = "spearman")

Blk.cor <- cor.test(ord_plot.data[,"Axis.2"], ord_plot.data[,"Db3rdbar"], method = "spearman")

Lat.cor <- cor.test(ord_plot.data[,"Axis.1"], ord_plot.data[,"Lat"], method = "spearman")
Long.cor <- cor.test(ord_plot.data[,"Axis.2"], ord_plot.data[,"Long"], method = "spearman")

#Plotting ordination 
Prec <- ggplot(ord_plot.data, aes(x = Axis.1, y = Precip_AMJ)) + 
  geom_smooth(method = lm, color = "gray14") +
  geom_point(aes(color = Lat), size = 3) + 
  theme_gray() +
  scale_color_gradientn(colours= c("#ef8a62","#67a9cf")) +
  labs(y = "Precipitation season (mm)", color = "Latitude") +
  annotate("text", x = 0.15, y = 210, 
           label = paste("p-value =", format(P.cor$p.value, digits = 3),
                         "\nrho =", round(P.cor$estimate, digits = 3)),
                         fontface = "bold", hjust = 0)


Tmin <- ggplot(ord_plot.data, aes(x = Axis.1, y = Tmin_AMJ)) + 
  geom_smooth(method = lm, color = "gray14") +
  geom_point(aes(color = Lat), size = 3) + 
  theme_gray() +
  scale_color_gradientn(colours= c("#ef8a62","#67a9cf")) +
  labs(y = "Minimum temperature season", color = "Latitude") +
  annotate("text", x = 0.15, y = 16, 
           label = paste("p-value =", format(Tmin.cor$p.value, digits = 3,
                                             scientific = TRUE),
                         "\nrho =", round(Tmin.cor$estimate, digits = 3)),
                         fontface = "bold", hjust = 0)

Clay <- ggplot(ord_plot.data, aes(x = Clay, y = Axis.2)) + 
  geom_smooth(method = lm, color = "gray14") +
  geom_point(aes(color = WC3rdbar), size = 3) + 
  theme_gray() +
  scale_color_gradientn(colours= c("#ef8a62","#67a9cf")) +
  labs(x = "Clay content (%)", color = "Vol. water\ncontent (%)") +
  annotate("text", x = 50, y = 0.38, 
           label = paste("p-value =", format(Clay.cor$p.value, digits = 3, 
                                             scientific = TRUE),
                         "\nrho =", round(Clay.cor$estimate, digits = 3)),
                         fontface = "bold", hjust = 0)

Blk <- ggplot(ord_plot.data, aes(x = Db3rdbar, y = Axis.2)) + 
  geom_smooth(method = lm, color = "gray14") +
  geom_point(aes(color = WC3rdbar), size = 3) + 
  theme_gray() +
  scale_color_gradientn(colours= c("#ef8a62","#67a9cf")) +
  labs(x = "Bulk density (g/cm3)", color = "Vol. water\ncontent (%)") +
  annotate("text", x = 1.58, y = 0.4, 
           label = paste("p-value =", format(Blk.cor$p.value, digits = 3,
                                             scientific = TRUE),
                         "\nrho =", round(Blk.cor$estimate, digits = 3)),
                         fontface = "bold", hjust = 0)

#Grid arregement for linear graphs
plot_grid(Prec, Tmin, Clay, Blk, labels = c("A", "B", "C", "D"), 
          ncol = 2, nrow = 2, align = "h")
```


## Mantel test for parameters
```{r Mantel_test}
Oom.dist <- distance(Oom_biom2, "bray")
env2 <- data.frame(env[,-c(1:7,12,24)])
env2.dist <- vegdist(env2$Lat, method = "euclidean")

# test.m <- mantel(Oom.dist, env2.dist, method = "pearson")
# test.m$statistic
# 
# length(colnames(env2))

factor_mantel <- function(dist, env){
  n <- length(colnames(env))
  df <- data.frame(Env.var = character(0), stat = numeric(0), pval = numeric(0))
  for (i in seq(1,n,1)) {
    factor_dist <- vegdist(env[,i], method = "euclidean")
    mt_test <- mantel(dist, factor_dist, method = "spearman")
    df <- rbind(df, data.frame(Env.var = colnames(env[i]), 
                               Statistic = mt_test$statistic, 
                               p.val = mt_test$signif))
    }
  df
}


mantel_table <- factor_mantel(Oom.dist, env2)
cor.table <- left_join(fit_data, mantel_table, by = "Env.var")
kable(cor.table, digits = 3)
```


## Abundance of top 8 pathogenic species
```{r map_pie_chart, fig.align='center', fig.width=11, fig.height=7}
##Top 8 species
top.sp <- names(sort(taxa_sums(Oom_phylo), TRUE)[1:8])
oom.sp <- prune_taxa(top.sp, Oom_phylo)
oom.sp.st <- merge_samples(oom.sp, "State2")

##Function
getLabelPoint <- function(county) {Polygon(county[c('long', 'lat')])@labpt}

#map data
library(maps)
library(sp)
states <- map_data("state")
centroids <- by(states, states$region, getLabelPoint)
centroids <- do.call("rbind.data.frame", centroids)
names(centroids) <- c('long', 'lat')

centroids[row.names(centroids) == "michigan",]$long <- -84.62014
centroids[row.names(centroids) == "michigan",]$lat <- 43.49422

#Transform counts for plot
oom.data.pie <- transform_sample_counts(oom.sp.st, 
                                           function(x) 100 * x/sum(x))
oom.data.pie <- psmelt(oom.data.pie)
oom.data.pie <- oom.data.pie[,c(1:3,6,55,56)] %>%
                mutate(sample = tolower(Sample))

oom.data.pie$sample <- gsub("n dakota","north dakota", oom.data.pie$sample)
oom.data.pie$sample <- gsub("s dakota","south dakota", oom.data.pie$sample)
oom.data.pie$Species <- factor(oom.data.pie$Species,
                               levels(oom.data.pie$Species)[c(7,3,6,8,1,2,5,4)])
#data
oom.pie.0 <- add_rownames(centroids, "sample") %>%
  left_join({distinct(oom.data.pie)}) %>%
  filter(!is.na(Clade))
  
oom.pie <- oom.pie.0 %>% split(., .$sample)

#Color scale
pal2 <- colorRampPalette(brewer.pal(8, "Set3"))

#Pie chart
pies <- setNames(lapply(1:length(oom.pie), function(i){
  ggplot(oom.pie[[i]], aes(x=1, Abundance, fill=Species)) +
    geom_bar(stat="identity", width=1, color="black") + 
    coord_polar(theta="y") + 
    theme_tree() + 
    xlab(NULL) + 
    ylab(NULL) + 
    theme_transparent() +
    scale_fill_manual(values = pal2(8)) +
    theme(plot.margin=unit(c(0,0,0,0),"mm"))
}), names(oom.pie))

#Legend
e1 <- ggplot(oom.pie[[2]], aes(x=1, Abundance, fill=Species)) +
        geom_bar(stat="identity", width=1) + 
        coord_polar(theta="y") + 
  scale_fill_manual(values = pal2(8), 
                    labels = c("Py. sylvaticum","Py. heterothallicum", "Py. oopapillum",
                               "Py. ultimum var. ultimum","Py. aff. dissotocum",
                               "Py. aff. torulosum", "Py. lutarium","Py. irregulare")) +
  theme(legend.text = element_text(face = "italic"))

leg1 <- gtable_filter(ggplot_gtable(ggplot_build(e1)), "guide-box") 

#map
states2 <- subset(states, region %in% c(unique(oom.data.pie$sample), "missouri"))
map.p <- ggplot(states2, aes(long, lat, group=group)) +  
    geom_polygon(fill="gray90", color = "gray20", size=0.125) +
    xlim(-105,-78) +
    theme_transparent(axis.title = element_blank(),
                      axis.text = element_blank(),
                      axis.line = element_blank(),
                      axis.ticks = element_blank()) +
      annotation_custom(grob = leg1, xmin = -81, xmax = -83, ymin = 33, ymax = 40) 

#Final plot
n <- length(pies)
for (i in 1:n) {
    nms <- names(pies)[i]
    dat <- oom.pie.0[which(oom.pie.0$sample == nms)[1], ]
    map.p <- subview(map.p, pies[[i]], x = unlist(dat[["long"]])[1], y=unlist(dat[["lat"]])[1], 0.07, 0.07)
}

print(map.p)
```

## Correlation of individual species with different parameters using occupancy models

```{r var_plots, warning=FALSE, message=FALSE, fig.height=7, fig.width=6}
library(ggradar)
library(scales)
library(MASS)
#Oom_phylo.sp <- transform_sample_counts(Oom_phylo, 
#                                     function(x) x/sum(x))
test <- psmelt(Oom_phylo)
head(test)
test %>% filter(Species == "Pythium_sylvaticum"| 
                Species == "Pythium_heterothallicum"|
                Species == "Pythium_oopapillum"|
                Species == "Pythium_ultimum_var._ultimum"|
                Species == "Pythium_aff._dissotocum"|
                Species == "Pythium_aff._torulosum") %>% 
  group_by(Species) %>%
  filter(!is.na(Clay)) %>%
  filter(CEC7 < 100) -> test_dat


pp1 <- ggplot(test_dat, aes(x=pHwater, y=log10(Abundance + 1))) + 
  geom_point(aes(colour=Lat), size = 3, alpha=0.9) + 
  geom_smooth(se = FALSE, method = "glm.nb", colour="grey30") +
  facet_wrap(~ Species) + labs(x="soil pH") +   
  scale_color_continuous(name="Latitude", na.value = "grey50", 
                         low = "#ef8a62", high = "#67a9cf") +
  theme_bw()

pp2 <-ggplot(test_dat, aes(x=Clay, y=log10(Abundance + 1))) + 
  geom_point(aes(colour=Lat), size = 3, alpha=0.9) + 
  geom_smooth(se = FALSE, method = "glm.nb", colour="grey30") +
  facet_wrap(~ Species) + labs(x="Clay percent (%)") +   
  scale_color_continuous(name="Latitude", na.value = "grey50", 
                         low = "#ef8a62", high = "#67a9cf") +
  theme_bw()

pp3 <-ggplot(test_dat, aes(x=Precip_AMJ, y=log10(Abundance + 1))) + 
  geom_point(aes(colour=Lat), size = 3, alpha=0.8) + 
  geom_smooth(se = FALSE, method = "glm.nb", colour="grey30") +
  facet_wrap(~ Species) + labs(x="Seasonal precipitation (mm)") +   
  scale_color_continuous(name="Latitude", na.value = "grey50", 
                         low = "#ef8a62", high = "#67a9cf") +
  theme_bw()

pp4 <- ggplot(test_dat, aes(x=Tmin_AMJ, y=log10(Abundance + 1))) + 
  geom_point(aes(colour=Lat), size = 3, alpha=0.9) + 
  geom_smooth(se = FALSE, method = "glm.nb", colour="grey30") +
  facet_wrap(~ Species) + labs(x="Season minimum temperature (ºC)") +   
  scale_color_continuous(name="Latitude", na.value = "grey50", 
                         low = "#ef8a62", high = "#67a9cf") +
  expand_limits(x=c(4,18)) +
  theme_bw()


pp5 <- ggplot(test_dat[test_dat$CEC < 100,], aes(x=CEC7, y=log10(Abundance + 1))) + 
  geom_point(aes(colour=Lat), size = 3, alpha=0.9) + 
  geom_smooth(se = FALSE, method = "glm.nb", colour="grey30") +
  facet_wrap(~ Species) + labs(x="Cation exchange capacity (meq/100g)") +   
  scale_color_continuous(name="Latitude", na.value = "grey50", 
                         low = "#ef8a62", high = "#67a9cf") +
  theme_bw()

plot_grid(pp1, pp2, labels = c("A", "B"), 
          ncol = 1, nrow = 2, align = "h")

plot_grid(pp3, pp4, labels = c("C", "D"), 
          ncol = 1, nrow = 2, align = "h")

plot_grid(pp5, labels ="E", 
          ncol = 1, nrow = 2, align = "h")
```

##Ordination by taxa

```{r Ordination_plot_taxa, fig.align='center', fig.width=19, fig.height=17}
Oom_biom2 <- prune_samples(!is.na(Oom_biom@sam_data$Lat), Oom_biom)

Oom_biom_ord <- ordinate(Oom_biom2, "PCoA", "bray")
ord_plot2 <- plot_ordination(Oom_biom2, Oom_biom_ord, type = "taxa", color = "Clade")
ord_plot.f2 <- ord_plot2 + geom_point(size = 4, alpha = 0.7) +
  theme_light() + labs(color = "Clade")

## Environment fit analysis
bray.pcoa2 <- ordinate(Oom_biom2, method = "PCoA", "bray")
#env3 <- vegansam(Oom_biom2) %>% select("CDL_2011":"Precip_AMJ")

Oom_env2 <- envfit(bray.pcoa$vectors, env, permutations = 999)

fit_data2 <- as.data.frame(scores(Oom_env2, display = "vectors")) %>%
  add_rownames(var = "Env.var") %>%
  bind_cols(data.frame(Oom_env2$vectors$r, Oom_env2$vectors$pvals)) %>%
  rename(R2 = Oom_env2.vectors.r, P.value = Oom_env2.vectors.pvals) %>%
  arrange(P.value)
  

## Vectors for plot
fit_reduced2 <- fit_data2[fit_data2$P.value < 0.05,] 

fit_plot2 <- as.data.frame(scores(Oom_env2, display = "vectors")) %>%
  add_rownames(var = "Env.var") %>%
  inner_join(fit_reduced, by = "Env.var") 


ord_plot.data2 <- plot_ordination(Oom_biom2, Oom_biom_ord, 
                                  type = "taxa", color = "Clade", justDF = TRUE)

(ord.plot.env2 <- ggplot(data = ord_plot.data2, aes(x = Axis.1, y = Axis.2)) + 
  labs(color = "Clade", x = "PCoA 1 [12.1%]", y = "PCoA 2 [9.4%]") +
  geom_segment(data = fit_plot2, aes(x = 0, xend = Axis.1.x, y = 0, yend = Axis.2.x), 
               arrow = arrow(length = unit(0.1,"cm")), color = "black", size = 0.8) + 
  geom_label_repel(data = fit_plot2, aes(x = Axis.1.x, y = Axis.2.x, label = Env.var), 
            size = 2, force = 1, label.size = 0.15) + 
  geom_point(aes(color=Clade), size = 4, alpha = 0.7) + 
  facet_wrap(~Clade) +
  theme_gray())
```
