---
title: "Oomycete species associated with soybean seedlings in the U.S."
author: "J. Alejandro Rojas"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
#Knitr options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

##Library load
# ipak function: install and load multiple R packages.
# check to see if packages are installed. Install them if they are not, then load them into the R session.
# Source: https://gist.github.com/stevenworthington/3178163
ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2","RColorBrewer","grid","gridExtra","plyr","lme4",
              "lsmeans","knitr","tidyr","dplyr", "MASS", "magrittr", 
              "reshape2", "FactoMineR", "cowplot")
ipak(packages)

##Other functions
#Standard error function
std <- function(x) sd(x)/sqrt(length(x))
CVcoeff <- function(x) (sd(x)/mean(x))*100

#Trimming leading or trailing space
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```

#Part I: identification and pathogenicity characterization

## Figures and analyses

__Figure 1.__ Map of sampled soybean fields in 2011 and 2012, and intensity of 
planted soybean acres demonstrated by color intensity at the county/parish level. 

```{r map_soybean_samples, echo=FALSE, fig.height=8, fig.width=12, cache=TRUE}
source("../code/soy_sample_map.R")

print(gg2)
print(gg3, vp=viewport(.65, .88, .4,.4))
```


__Figure 2.__ Frequency at which different Oomycete species were recovered from 
diseased soybean seedlings in 2011 and 2012.  (*) Species previously reported as 
associated with soybean in the fungal-host database (<http://nt.ars-grin.gov/fungaldatabases/fungushost/fungushost.cfm>).
```{r abundance_plot, fig.height=10, fig.width=8, echo=FALSE, fig.align='center'}
source("../code/abundance_analysis.R")

#Plots
Isol_11 <- ggplot(data=Data_11_12, aes(x=reorder(Species, freq_11, mean), y=freq_11)) + 
  geom_bar(stat="identity", color="#636363", fill="#7fbf7b") + 
  theme_bw() +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(),
        plot.margin = unit(c(1,-4,1,1), "mm")) +
  labs(y="Frequency (%)") +  xlab(NULL) +
  coord_flip() + scale_y_reverse()

Isol_12 <- ggplot(data=Data_11_12, aes(x=reorder(Species, freq_11, mean), y=freq_12)) + 
  geom_bar(stat="identity", color="#636363", fill="#9ecae1") + 
  xlab(NULL) + theme_bw() + 
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.margin = unit(c(1,1,1,-4), "mm")) +
    labs(y="Frequency (%)") + coord_flip()

mid.Isol <- ggplot(data=Data_11_12, aes(x=reorder(Species, freq_11, mean), y=0)) +
  geom_bar(stat="identity",fill="white", alpha=0) + 
  geom_text(aes(y=0, label=Species), fontface="italic", size=3) +
  geom_segment(aes(y=-0.003,yend=-0.0025,xend=Species), size=0.2) +
  geom_segment(aes(y=0.0025,yend=0.003,xend=Species), size=0.2) +
  xlab(NULL) +
  coord_flip() + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.background=element_blank(),
        panel.grid=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_text(color=NA),
        axis.ticks.x=element_line(color=NA),
        axis.title.x=element_text(color=NA),
        plot.margin = unit(c(1,-4,1,-6.5), "mm"))

#Final abundance plot
grid.arrange(ggplotGrob(Isol_11), ggplotGrob(mid.Isol), ggplotGrob(Isol_12), ncol=3, widths=c(.3,.2,.3))
```

__Figure 3.__  Frequency of the oomycete species summarized by clade for 2011 
and 2012. Oomycete genera outside of Pythium and Phytophthora were summarized 
by genus.  Those species designated as spp. are not well resolved based only on
the ITS sequence.
```{r clade_plot, fig.height=5, fig.width=7, echo=FALSE, fig.align='center'}
#Summarise data by year
Clade_data11 <- ddply(Isolate_11, "Clade", summarise,
                   N = as.numeric(length(qDef)),
                   freq = (N/length(Isolate_11$Year))*100
)

Clade_data12 <- ddply(Isolate_12, "Clade", summarise,
                   N = as.numeric(length(qDef)),
                   freq = (N/length(Isolate_12$Year))*100
)
#Re-merge summarized data by year and replace NAs by 0
Clade_11_12 <- full_join(Clade_data12, Clade_data11, by="Clade") %>%
  dplyr::rename(N_12=N.x, y2012=freq.x, N_11=N.y, y2011=freq.y)

Clade_11_12[is.na(Clade_11_12)] <- 0

Clade_data <-  tidyr::gather(Clade_11_12, Year, Freq, c(y2012, y2011))

Clade_data$Year <- factor(Clade_data$Year, levels = c("y2012","y2011"))
Clade_data$Clade <- factor(Clade_data$Clade,
                               levels(Clade_data$Clade)[c(1,5,7,13,8,
                                                          2,4,6,10,16,
                                                          11,3,15,14,9,12)])
#Plots
(Clade_plot <- ggplot(data=Clade_data, aes(x=Clade, y=Freq)) + 
  geom_bar(aes(fill=Year), stat="identity", position="dodge", colour="#636363") + 
  theme_bw() +
  scale_fill_manual(values=c("#9ecae1","#7fbf7b"), labels=c("2012","2011"),
                    guide = guide_legend(reverse = TRUE)) +
  theme(axis.text.y=element_text(face="italic", size = 12)) +
  labs(y="Frequency (%)") + xlab(NULL) + coord_flip())
```

For __seed rot analyses__ in detail, please check the [seed_rot_data_analysis.Rmd](Seed_rot_data_analysis.md)__

__Figure 3.__
```{}

```

## Tables

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
