---
title: "Seedling pathogenicity data analysis"
author: "Alejandro Rojas"
date: "September 23, 2015"
output: 
  html_document: 
    keep_md: yes
---

#Pathogenicity data analysis - Soybean oomycete pathogen survey

```{r custom_functions, echo=FALSE, message=FALSE}
#Standard error function
std <- function(x) sd(x)/sqrt(length(x))
CVcoeff <- function(x) (sd(x)/mean(x))*100

# ipak function: install and load multiple R packages.
# check to see if packages are installed. Install them if they are not, then load them into the R session.
# Source: https://gist.github.com/stevenworthington/3178163
ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
sapply(pkg, require, character.only = TRUE)
}

#Trimming leading or trailing space
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```

```{r libraries, echo=FALSE, message=FALSE}
packages <- c("ggplot2","RColorBrewer","grid","gridExtra","plyr","lme4","lsmeans","knitr","tidyr","dplyr", "MASS", "magrittr", "reshape2", "FactoMineR", "cowplot")
ipak(packages)

#require(devtools)
#install_github('vqv/ggbiplot')
```

##Seed dry weight - Pathogenicity data
Processing the data for graphical exploration and analysis.  
```{r file_reading}
#Reading the file
dry_weight <- read.csv(file = "../data/clean/dry_weight_allsets.csv")
dry_weight$Species <- trim(dry_weight$Species)

#Calculate root:shoot ratio
dry_weight$r_s <- with(dry_weight, Weight.per.root/Weight.per.shoot)

#Summarizing data using different parameters by plyr library
dryw_sum <- ddply(dry_weight, c("Species"), summarise,
              N = length(Weight.per.root),
              mean_wpr = mean(Weight.per.root), 
              sd_wpr = sd(Weight.per.root),
              se_wpr = sd_wpr/sqrt(N),
              mean_wps = mean(Weight.per.shoot), 
              sd_wps = sd(Weight.per.shoot),
              se_wps = sd_wps/sqrt(N),
              mean_rs = mean(r_s),
              sd_rs = sd(r_s),
              se_rs = sd_rs/sqrt(N)
              )
  
#Setting limits for error bars
wpr_limits <- aes(ymax = mean_wpr + se_wpr, ymin=mean_wpr - se_wpr)
wps_limits <- aes(ymax = mean_wps + se_wps, ymin=mean_wps - se_wps)
rs_limits <- aes(ymax = mean_rs + se_rs, ymin=mean_rs - se_rs)

#melt data into long format
dryw_sum.t1 <- melt(dryw_sum, measure.vars = c("mean_wpr","mean_wps","mean_rs")) %>%
  subset(select = c(Species, variable,value))

dryw_sum.t2 <- melt(dryw_sum, measure.vars = c("se_wpr","se_wps","se_rs")) %>% 
  subset(select = c(Species, variable,value)) %>%
  dplyr::rename(Spp=Species, SE_var=variable, SE_val=value)

dryw_sum.t <- cbind(dryw_sum.t1, dryw_sum.t2)
```

Plots for dry weight per root, dry weight per shoot and root:shoot ratio.
```{r point_weight, fig.align='center', fig.width=11, fig.height=8}
plot_wpr <- ggplot(dryw_sum, aes(x = reorder(Species, mean_wpr, median), y = mean_wpr)) +
  geom_point(stat = "summary", fun.y = "mean", size = 3, col="#d8b365") +
  geom_errorbar(wpr_limits, width=0.2, col="#d8b365") + theme_gray() +
  theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, face="italic"), 
        axis.text.y=element_text(angle=90, hjust = 0.5), 
        axis.title.x = element_text(angle=180),
        plot.margin=unit(c(1,1,1,1), "mm")) +
  labs(x="Species", y = "Weight per root (mg)") 

plot_wps <- ggplot(dryw_sum, aes(x = reorder(Species, mean_wpr, median), y = mean_wps)) +
  geom_point(stat = "summary", fun.y = "mean", size = 3, col="#7fbf7b") +
  geom_errorbar(wps_limits, width=0.2, col="#7fbf7b") + theme_gray() +
  #theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, face="italic"))) +
  theme(axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.text.y=element_text(angle=90, hjust = 0.5),
        plot.margin=unit(c(1,1,-7,1), "mm")) +
  labs(x="", y = "Weight per shoot (mg)") 

plot_rs <- ggplot(dryw_sum, aes(x = reorder(Species, mean_wpr, median), y = mean_rs)) +
  geom_point(stat = "summary", fun.y = "mean", size = 3, col="#4d4d4d") +
  geom_errorbar(rs_limits, width=0.2, col="#4d4d4d") + theme_gray() +
  #theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, face="italic"))) +
  theme(axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.text.y=element_text(angle=90, hjust = 0.5),
        plot.margin=unit(c(1,1,-7,1), "mm")) + 
  labs(x="", y = "root:shoot ratio") 

grid.draw(rbind(ggplotGrob(plot_rs), ggplotGrob(plot_wps), ggplotGrob(plot_wpr), size="last"))

```


##Root measurements - pathogenicity data
```{r data_input}
#Reading the file
root_raw <- read.csv(file = "../data/clean/root_measurements_final.csv")
root_raw$Species <- trim(root_raw$Species)

#Data transformation using different parameters by plyr library
root_measures <- ddply(root_raw, c("Species","Isolate","Set","Rep"), summarise,
              N = length(Area_cm),
              area.r = mean(Area_cm), 
              length.r = mean(Length_cm)
              )

#Summarizing data using different parameters by plyr library
root_sum <- ddply(root_measures, c("Species"), summarise,
              N = length(area.r),
              area.p = mean(area.r),
              sd_ap = sd(area.r),
              se_ap = sd_ap/sqrt(N),
              ln.p = mean(length.r),
              sd_lp = sd(length.r),
              se_lp = sd_lp/sqrt(N)
              )

#Setting limits for error bars
ap_limits <- aes(ymax = area.p + se_ap, ymin=area.p - se_ap)
lp_limits <- aes(ymax = ln.p + se_lp, ymin=ln.p - se_lp)
```

Plots for the root area and root length
```{r point_scans, fig.align='center', fig.width=11, fig.height=8}
plot_area <- ggplot(root_sum, aes(x=reorder(Species, area.p, median), y=area.p)) +
  geom_point(stat = "summary", fun.y = "mean", size=3, col="#4d4d4d") +
  geom_errorbar(ap_limits, width=0.2, col="#4d4d4d") + theme_gray() +
  theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, face="italic"),
        axis.text.y=element_text(angle=90, hjust = 0.5),
        plot.margin=unit(c(1,1,1,1), "mm")) +
  labs(x="Species", y = expression(paste("Root area (",cm^"2",")")))

plot_len <- ggplot(root_sum, aes(x=reorder(Species, area.p, median), y=ln.p)) +
  geom_point(stat = "summary", fun.y = "mean", size=3, col="#7fbf7b") +
  geom_errorbar(lp_limits, width=0.2, col="#7fbf7b") + theme_gray() +
  #theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, face="italic"),
  #      axis.text.y=element_text(angle=90)) +
  theme(axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.text.y=element_text(angle=90, hjust = 0.5),
        plot.margin=unit(c(1,1,-7,1), "mm")) +
  labs(x="", y = "Root Length (cm)")

grid.draw(rbind(ggplotGrob(plot_len), ggplotGrob(plot_area), size="last"))
```

Merge data for root weight, shoot weight, and root measurements.
```{r combined_data}
#Data for table publication
Root_data <- left_join(dryw_sum,root_sum, by="Species")
Root_data$N.x <- with(Root_data, N.x/9)

#Selecting the columns desired
Root_data <- Root_data %>% dplyr::select(Species,N.x,mean_wpr,
                                         se_wpr,mean_wps,se_wps,
                                         mean_rs,se_rs, area.p,se_ap,ln.p,se_lp)

#Renaming column data
Root_data <- dplyr::rename(Root_data, N=N.x, Weight.Root=mean_wpr, se.wr=se_wpr,
                    Weight.Shoot=mean_wps, se.ws=se_wps, Root.Shoot=mean_rs,
                    se.rs=se_rs, Root.Area=area.p, se.area=se_ap,
                    Root.length=ln.p, se.length=se_lp)

kable(Root_data, digits = 3, format = "markdown")
```

## PCA analysis
Evaluation of data and factor mining
```{r PCA_analysis}
## Combine dataset for multivariate analysis
root_final <- full_join(root_measures,dry_weight, by=c("Species","Isolate","Set","Rep"))
root_final <- root_final[c(-5,-8,-9,-10)]

## Log transforming data
root.log <- root_final
root.log[,5:9] <- log10(root.log[,5:9] + 1)

#### PCA exploration of the data 
library(FactoMineR)
rd.pca <- PCA(root_final[5:9], scale.unit = TRUE, ncp=5, graph = T)

rd.pca$var$contrib
dimdesc(rd.pca, axes = c(1,2))

rd.hcpc <- HCPC(rd.pca, nb.clust = 3)
rd.hcpc$desc.var

root.clusters <- rd.hcpc$data.clust$clust

root.log.pca <- prcomp(root_final[5:9],
                       center = TRUE,
                       scale. = TRUE)
summary(root.log.pca)

#Bi-plot for PCA analysis
library(ggbiplot)
g <- ggbiplot(root.log.pca, obs.scale = 1, var.scale = 1, 
              groups = root.clusters, ellipse = TRUE, 
              circle = FALSE, alpha=0.2)
g <- g + scale_color_discrete(name = '') + theme_gray() 
(g <- g + theme(legend.direction = 'horizontal', 
               legend.position = 'top'))
```


## MANOVA analysis
```{r MANOVA analysis}
## MANOVA analysis
# fit <- manova(cbind(area.r, length.r, Weight.per.root)~Species, data=root_scale)
# summary(fit, test=c("Pillai"))
# summary(fit, test=c("Wilks"))
# summary(fit, test=c("Hotelling-Lawley"))
library(nlme)
fit <- lme(cbind(area.r, length.r, Weight.per.root) ~ Species, random= ~1|Isolate, data=root.log)
fit1 <- lme(cbind(area.r, length.r, Weight.per.root) ~ Species, random= ~1|Set/Isolate, data=root.log)
fit2 <- lme(cbind(area.r, length.r, Weight.per.root) ~ Species, random = ~1|Set, data=root.log)
fit3 <- lme(cbind(area.r, length.r, Weight.per.root) ~ Species,
           random = ~1|Set/Species/Isolate, data=root.log)

#model tests
anova(fit,fit1,fit2, fit3)

#Checking correlation of dependent variables
library(car)
scatterplotMatrix(~ area.r + length.r + Weight.per.root, data = root.log)

cor.test(root.log$area.r, root.log$length.r, method="pearson")
cor.test(root.log$Weight.per.root, root.log$length.r, method="pearson")
cor.test(root.log$area.r, root.log$Weight.per.root, method="pearson")

cor.test(root_final$area.r, root_final$length.r, method="pearson")
cor.test(root_final$Weight.per.root, root_final$length.r, method="pearson")
cor.test(root_final$area.r, root_final$Weight.per.root, method="pearson")

library(lsmeans)
plot(fit1)
#Variance
VarCorr(fit1)

#Confidence intervals
fit1.ci <- intervals(fit1)
fit1.ci.2 <- as.data.frame(fit1.ci[[1]]) %>% add_rownames(var = "Species") %>% arrange(est.)
ggplot(fit1.ci.2, aes(x = reorder(Species, est., mean), 
                 y = est., ymin = lower, ymax = upper)) + 
  geom_errorbar() + geom_point() + coord_flip() + theme_bw()

#Contrasts
fit1.ls <- lsmeans(fit1, "Species")
plot(fit1.ls)
fit1.d <- contrast(fit1.ls, "trt.vs.ctrl", ref=4, adjust="bon")
fit1.dc <- as.data.frame(summary(fit1.d))
kable(fit1.dc, format = "markdown")

#### Univariate analysis
#### Area
fit.area <- lme(area.r ~ Species, random= ~1|Set/Isolate, data=root.log)
fit.area.ls <- lsmeans(fit.area, "Species")
fit.area.d <- contrast(fit.area.ls, "trt.vs.ctrl", ref=4, adjust="bon")
area.d.t <- as.data.frame(summary(fit.area.d))
#### Length
fit.length <- lme(length.r ~ Species, random= ~1|Set/Isolate, data=root.log)
fit.length.ls <- lsmeans(fit.length, "Species")
fit.length.d <- contrast(fit.length.ls, "trt.vs.ctrl", ref=4, adjust="bon")
length.d.t <- as.data.frame(summary(fit.length.d))
#### Weight per root
fit.wpr <- lme(Weight.per.root ~ Species, random= ~1|Set/Isolate, data=root.log)
fit.wpr.ls <- lsmeans(fit.wpr, "Species")
fit.wpr.d <- contrast(fit.wpr.ls, "trt.vs.ctrl", ref=4, adjust="bon")
wpr.d.t <- as.data.frame(summary(fit.wpr.d))

#### Merge results
univ.t1 <- full_join(area.d.t, length.d.t, by="contrast") %>% 
  dplyr::select(c(contrast, contains("p.value"))) %>%
  dplyr::rename(ar.p = p.value.x, ln.p = p.value.y)

univ.t2 <- full_join(univ.t1, wpr.d.t, by="contrast") %>%
  dplyr::select(c(contrast, ar.p, ln.p, p.value)) %>%
  dplyr::rename(wpr.p = p.value)

univ.t <- full_join(univ.t2, fit1.dc, by="contrast") %>%
  dplyr::select(c(contrast, p.value, ar.p, ln.p, wpr.p)) %>%
  dplyr::rename(mnv.p = p.value)

univ.t <- cbind(colsplit(univ.t$contrast," - ", names=c('Species','contrast')),univ.t[2:5])

kable(univ.t, digits = 3, format = "markdown")
```

```{r box_plot_trt}
### Creating two data sets for downstream analysis
#### Root original data set
root_final$Treatment[root_final$Species=="CONTROL"] <- "Control"
root_final$Treatment[root_final$Species=="CONTROL_NR"] <- "Control"
root_final$Treatment[is.na(root_final$Treatment)] <- "Inoculated"
rownames(root_final) <- make.names(root_final[,1], unique = TRUE)

t.test(root_final$Weight.per.shoot~root_final$Treatment)
t.test(root_final$Weight.per.root~root_final$Treatment)
t.test(root_final$length.r~root_final$Treatment)
t.test(root_final$area.r~root_final$Treatment)

## Analyzing by variable
##Function boxplot
draw_bp <- function(d, v, c, l){
  plot1 <- ggplot(data=d, aes_string(y=v, x=c)) +
    geom_boxplot(position="identity") + theme_bw() + labs(y=l)
  plot1
}

ar.root <- draw_bp(root_final, "area.r", "Treatment", expression(paste("Root area (",cm^"2",")")))
ln.root <- draw_bp(root_final, "length.r", "Treatment", "Root length (cm)")
wpr <- draw_bp(root_final, "Weight.per.root", "Treatment", "Weight per root (mg)")
wpp <- draw_bp(root_final, "Weight.per.shoot", "Treatment", "Weight per shoot (mg)")


plot_grid(ar.root,ln.root,wpr,wpp, ncol=2, nrow=2, labels = c("A","B","C","D"))
```



```{r data_plot_final}
## Summarizing data by species - full dataset
root.sp <- ddply(root_final, c("Species"), summarise,
              N = length(area.r),
              ar = mean(area.r), 
              ar.se = sd(area.r)/sqrt(N),
              ln = mean(length.r),
              ln.se = sd(length.r)/sqrt(N),
              wpr = mean(Weight.per.root),
              wpr.se = sd(Weight.per.root)/sqrt(N)
              )

root.sp <- dplyr::full_join(root.sp, univ.t, by="Species")
root.sp[is.na(root.sp)] <- 1

kable(root.sp, digits = 3, format = "markdown")

root.sp$ar.sg <- ifelse(root.sp$ar.p<0.05,"SG","NS")
root.sp$ln.sg <- ifelse(root.sp$ln.p<0.05,"SG","NS")
root.sp$wpr.sg <- ifelse(root.sp$wpr.p<0.05,"SG","NS")
#Setting limits for error bars
ar_lt <- aes(ymax = ar + ar.se, ymin=ar - ar.se)
ln_lt <- aes(ymax = ln + ln.se, ymin=ln - ln.se)
wpr_lt <- aes(ymax = wpr + wpr.se, ymin=wpr - wpr.se)
```


```{r final_plot, fig.align='center', fig.width=11, fig.height=8}

plot_ar <- ggplot(root.sp, aes(x=reorder(Species, ar, median), y=ar)) +
  geom_point(aes(shape=ar.sg), stat = "summary", 
             fun.y = "mean", size=3, colour="#4d4d4d") +
  geom_errorbar(ar_lt, width=0.2, col="#4d4d4d")  + theme_gray() +
  scale_shape_manual(values = c(21,19)) +
  theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, 
                                 face="italic", size = 12),
        axis.text.y=element_text(angle=90, hjust = 0.5),
        plot.margin=unit(c(1,1,1,1), "mm"),
        legend.position="none") +
  labs(x="Species", y = expression(paste("Root area (",cm^"2",")")))

plot_ln <- ggplot(root.sp, aes(x=reorder(Species, ar, median), y=ln)) +
  geom_point(aes(shape=ln.sg), stat = "summary", fun.y = "mean", 
             size=3, colour="#005824") +
  geom_errorbar(ln_lt, width=0.2, col="#005824")  + theme_gray() + 
  scale_shape_manual(values = c(21,19)) +
   #theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, face="italic"),
  #      axis.text.y=element_text(angle=90)) +
  theme(axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.text.y=element_text(angle=90, hjust = 0.5),
        plot.margin=unit(c(1,1,-7,1), "mm"),
        legend.position="none") +
  labs(x="", y = "Root Length (cm)")

plot_wpr <- ggplot(root.sp, aes(x=reorder(Species, ar, median), y=wpr)) +
  geom_point(aes(shape=wpr.sg), stat = "summary", fun.y = "mean",
             size=3, colour = "#bf812d") +
  geom_errorbar(wpr_lt, width=0.2, col="#bf812d") + theme_gray() +
  scale_shape_manual(values = c(21,19)) +
  #theme(axis.text.x=element_text(angle=90, hjust = 1, vjust = 0.5, face="italic"),
  #      axis.text.y=element_text(angle=90)) +
  theme(axis.ticks = element_blank(), 
        axis.text.x = element_blank(),
        axis.text.y=element_text(angle=90, hjust = 0.5),
        plot.margin=unit(c(1,1,-7,1), "mm"),
        legend.position="none") +
  labs(x="", y = "Weight per root (mg)")

grid.draw(rbind(ggplotGrob(plot_wpr), ggplotGrob(plot_ln), ggplotGrob(plot_ar), size="last"))
```

## Boxplot based on significance
```{r fig.align='center', fig.height= 3, fig.width=10}
root.sp1 <- root.sp
root.sp1$Group[root.sp1$mnv.p>0.05] <- "Group 1"
root.sp1$Group[(root.sp1$mnv.p<0.05) & (root.sp1$ln.p>0.05) | (root.sp1$mnv.p<0.05) &(root.sp1$wpr.p>0.05)] <- "Group 2"
root.sp1$Group[(root.sp1$mnv.p<0.05) & (root.sp1$ln.p<0.05) & (root.sp1$wpr.p<0.05)] <- "Group 3"


## Plots by clusters
a <- draw_bp(root.sp1, "ar", "Group", expression(paste("Root area (",cm^"2",")")))
b <- draw_bp(root.sp1, "ln", "Group", "Root length (cm)")
c <- draw_bp(root.sp1, "wpr", "Group", "Weight per root (mg)")

plot_grid(a,b,c, ncol=3, nrow=1, labels = c("A","B","C"))

#function for actual mode of data
#mode_f <- function(x){as.integer(which.max(table(x)))}
```


### LDA Analysis (Not included)
```{r}
#### LDA analysis data set
root.lda.data <- cbind(root_final[1:4],log10(root_final[,5:9] + 1), root_final[10])
root.lda.data[5:8] <- scale(root.lda.data[5:8], center = TRUE, scale = TRUE)

root_lda <- lda(Species ~ area.r + length.r + Weight.per.root, data=root.lda.data)
root.lda.p <- predict(root_lda)

prop.lda <- root_lda$svd^2/sum(root_lda$svd^2)

dataset = data.frame(species = root.lda.data[,"Treatment"],
                     pca = root.log.pca$x, lda = root.lda.p$x)

(p1 <- ggplot(dataset) + geom_point(aes(lda.LD1, lda.LD2, colour = species, shape = species), size = 2.5, position = position_jitter(width=0.3, height=0.3)) + theme_gray() + 
  labs(x = paste("LD1 (", percent(prop.lda[1]), ")", sep=""),
       y = paste("LD2 (", percent(prop.lda[2]), ")", sep="")))
```


### Prevalence of pathogenic species
```{r}
#Vector for pathogenic species
c3 <- root.sp1[(root.sp1$mnv.p<0.05) & (root.sp1$ln.p<0.05),1]

#Read abundance data
Isolate_data <- read.csv("../data/clean/Isolates_11-12_final.csv")

#Summarise data by year
Data_11_12 <- ddply(Isolate_data, c("Species","State"), summarise,
                   N = as.numeric(length(qDef)),
                   freq = (N/length(Isolate_data$Year))*100
                   )

#Subsetting the data
Data_11_12.sub <- subset(Data_11_12, Data_11_12$Species %in% c3, c(Species, State, N))
D11_12 <- spread(Data_11_12.sub, State, N)
D11_12[is.na(D11_12)] <- 0
D11_12 <- gather(D11_12, State, N, Arkansas:Wisconsin)

D11_12$brk <- cut(D11_12$N, breaks = c(-0.5,0,10,50,100,150,200),
                  labels=c("0","1-10","10-50","50-100","100-150","150-200"))

#Tile plot
(prev.plot <- ggplot(D11_12, aes(x=State,y=reorder(Species, N, median))) + 
  geom_tile(aes(fill=brk)) + 
  scale_fill_brewer(palette = "GnBu", type = "seq", name="Isolates (n)") +
  theme_bw() +
  labs(y="Species") + 
  theme(axis.text.x=element_text(angle=-60, hjust=0, colour="black"), 
        axis.text.y=element_text(face="italic", colour="black")))

## Non pathogenic species prevalence
c4 <- root.sp1[(root.sp1$mnv.p>0.05) & (root.sp1$ln.p>0.05),1]

Data_11_12.sub2 <- subset(Data_11_12, Data_11_12$Species %in% c4, c(Species, State, N))
D11_12.2 <- spread(Data_11_12.sub2, State, N)
D11_12.2[is.na(D11_12.2)] <- 0
D11_12.2 <- gather(D11_12.2, State, N, Arkansas:Wisconsin)

D11_12.2$brk <- cut(D11_12.2$N, breaks = c(-0.5,0,10,50,100,150,200), 
                  labels=c("0","1-10","10-50","50-100","100-150","150-200"))

(prev.plot2 <- ggplot(D11_12.2, aes(x=State,y=reorder(Species, N, median))) + 
  geom_tile(aes(fill=brk)) + 
  scale_fill_brewer(palette = "GnBu", type = "seq", name="Isolates (n)") +
  theme_bw() +
  labs(y="Species") + 
  theme(axis.text.x=element_text(angle=-60, hjust=0, colour="black"), 
        axis.text.y=element_text(face="italic", colour="black")))
```
